### 2023-09-05 sections/1-introduction.tex
fix
fixed unfortunate typos

\section{Introduction}
\label{sec:introduction}
Often dubbed "the Big Data explosion," rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
This is especially true in the field of astronomy. 
The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions. 
Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.

Biology has also seen an influx of large, high-dimensional datasets. Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 
million genomes with aligned length of 50,000 letters; 
the GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length.

@@

For the sake of clarity, it is important to be precise about what we mean by \emph{exact} $k$-NN, and exact search in general.
In the context of $\rho$-NN search, exact search typically means returning only those data points that are identical to the query.
There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to $\rho$-NN search.
In contrast, in the context of $k$-NN search, while all approaches return exactly $k$ elements (ignoring how distance ties are handled), some approaches may miss some of the $k$ nearest elements, and erroneously include some of the elements outside the truly nearest $k$.
In any case, $k$-NN is typically used for inference; the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned; no closer points are elided and no further points are included.
This has implications for the accuracy of classification, but does not guarantee perfect classification accuracy.

There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance}; this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could be exhibit suboptimal accuracy.

@@

This paper focuses on four novel algorithms for exact $k$-NN search. We also present some 
improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, 
as well as improved genericity across distance functions. Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
Finally, we provide a comparison of 
CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets 
from the ANN benchmarks suite [CITE].

>>>>

\section{Introduction}
\label{sec:introduction}

Often dubbed ``the Big Data explosion,'' rapidly advancing technologies have enabled researchers to collect complex data at an unprecedented scale. 
This is especially true in the field of astronomy. 
The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss}, for example, in its APOGEE~\cite{alam2015eleventh} dataset, has collected near-infrared spectra of approximately 430,000 stars in approximately 9,000 dimensions. 
Their MaNGA dataset also contains spectral measurements across the face of each of about 10,000 galaxies across 2,700 deg$^2$.

Biology has also seen an influx of large, high-dimensional datasets.
The GreenGenes project~\cite{desantis2006greengenes} provides a multiple-sequence alignment of over one million bacterial 16S sequences, each 7,682 characters in length;
Silva 18S~\cite{10.1093/nar/gks1219} contains ribosomal DNA sequences of approximately 2.25 million genomes with aligned length of 50,000 letters.

@@

For the sake of clarity, it is important to be precise about what we mean by \emph{exact} $k$-NN, and exact search in general.
In the context of $\rho$-NN search, exact search typically means returning only those data points that are identical to the query.
There are many algorithms that perform well here, such as FM-Index~\cite{ferragina2000opportunistic}, which are less applicable to approximate search.
In contrast, in the context of $k$-NN search, while all approaches return exactly $k$ elements (ignoring how distance ties are handled), some approaches may miss some of the $k$ nearest elements, and erroneously include some of the elements outside the truly nearest $k$.
In any case, $k$-NN is commonly used for inference;
the \emph{prediction} of a class for a query is obviously not guaranteed to be correct.
When we refer to \emph{exact} $k$-NN, we refer to precisely the $k$ nearest neighbors being returned;
no closer points are elided and no further points are included.
This has implications for the accuracy of classification, but does not guarantee perfect classification accuracy.

There is some evidence that non-metrics such as cosine distance perform poorly for $k$-NN in biomedical settings~\cite{hu2016distance};
this may suggest that approximate $k$-NN (which may have imperfect precision and recall) could exhibit suboptimal accuracy.

@@

This paper focuses on four novel algorithms for exact $k$-NN search.
We also present some improvements to the clustering and $\rho$-nearest neighbors search algorithms in CHESS, as well as improved genericity across distance functions.
Additionally, we prove a guarantee about the scaling behavior of cluster radii. 
Finally, we provide a comparison of CAKES to the state-of-the-art in similarity search, FAISS~\cite{johnson2019billion} and HNSW, on several datasets from the ANN benchmarks suite~\cite{aumuller2020ann}.
